---
title: "Milestone Report"
author: "Gabriela Ochoa"
date: "06 aprile 2021"
output: html_document
---

# *Overview*

The goal of this project is just to display that you've gotten used to working with the data and that you are on track to create your prediction algorithm. Please submit a report on R Pubs (http://rpubs.com/) that explains your exploratory analysis and your goals for the eventual app and algorithm. This document should be concise and explain only the major features of the data you have identified and briefly summarize your plans for creating the prediction algorithm and Shiny app in a way that would be understandable to a non-data scientist manager. You should make use of tables and plots to illustrate important summaries of the data set. The motivation for this project is to: 1. Demonstrate that you've downloaded the data and have successfully loaded it in.2. Create a basic report of summary statistics about the data sets.3. Report any interesting findings that you amassed so far.4. Get feedback on your plans for creating a prediction algorithm and Shiny app. 
Review criteria

- Does the link lead to an HTML page describing the exploratory analysis of the training data set?
- Has the data scientist done basic summaries of the three files? Word counts, line counts and basic data tables?
- Has the data scientist made basic plots, such as histograms to illustrate features of the data?
- Was the report written in a brief, concise style, in a way that a non-data scientist manager could appreciate?## 

## _Sampling_

To reiterate, to build models you don't need to load in and use all of the data. Often relatively few randomly selected rows or chunks need to be included to get an accurate approximation to results that would be obtained using all the data. Remember your inference class and how a representative sample can be used to infer facts about a population.

To reiterate, to build models you don't need to load in and use all of the data. Often relatively few randomly selected rows or chunks need to be included to get an accurate approximation to results that would be obtained using all the data. Remember your inference class and how a representative sample can be used to infer facts about a population. You might want to create a separate sub-sample dataset by reading in a random subset of the original data and writing it out to a separate file. That way, you can store the sample and not have to recreate it every time. You can use the rbinom function to "flip a biased coin" to determine whether you sample a line of text or not.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### **Loading libraries**

```{r, results='hide'}
if(Sys.getenv("JAVA_HOME")!="")
      Sys.setenv(JAVA_HOME="")
library(rJava)
library(stringi)
library(openNLP)
library(qdap)
library(tm)
library(tokenizers)
library(NLP)
library(RWeka)
```

### **Loading data**

```{r}
# Read the blogs and twitter files using readLines
blog_data <- readLines("final/en_US/en_US.blogs.txt", warn = FALSE, encoding = "UTF-8")
twitter_data <- readLines("final/en_US/en_US.twitter.txt", warn = FALSE, encoding = "UTF-8")

# Read the news file using binary/binomial mode as there are special characters in the text
con <- file("final/en_US/en_US.news.txt", open="rb")
news_data <- readLines(con, encoding = "UTF-8")
close(con)
rm(con)
```

### **Calculating data dimensions** _Megabytes_

```{r}
## size of the data
blog_dim <- file.info("final/en_US/en_US.blogs.txt")$size / 1024 ^ 2
sprintf("The en_US.blogs.txt file is: %s Megabytes", blog_dim)

news_dim <- file.info("final/en_US/en_US.news.txt")$size / 1024 ^ 2
sprintf("The en_US.news.txt file is: %s Megabytes", news_dim)

twitter_dim <- file.info("final/en_US/en_US.twitter.txt")$size / 1024 ^ 2
sprintf("The en_US.twitter.txt file is: %s Megabytes", twitter_dim)

```

### **Data Summary**
```{r}
Data_info <- data.frame('File' = c("Blogs","News","Twitter"),
                      "FileSizeinMB" = c(blog_dim, news_dim, twitter_dim),
                      'NumberofLines' = sapply(list(blog_data, news_data, twitter_data), function(x){length(x)}),
                      'TotalCharacters' = sapply(list(blog_data, news_data, twitter_data), function(x){sum(nchar(x))}),
                      TotalWords = sapply(list(blog_data,news_data,twitter_data),stri_stats_latex)[4,],
                      'MaxCharacters' = sapply(list(blog_data, news_data, twitter_data), function(x){max(unlist(lapply(x, function(y) nchar(y))))})
                      )

Data_info

```
As we can see, each files has 200 & below `MB`and number of words are more than 30 million per file.  
- **Twitter** has the most amount of lines, and fewer words per line.   
- **Blogs** has the longest line (40,833 characters).  
- **News** has the longest paragraphs. 


### **Counting words**
In the `en_US twitter data set`, if you divide the number of lines where the word `"love"` (all lowercase) occurs by the number of lines the word `"hate"` (all lowercase) occurs, about what do you get?
```{r}
love_hate<-length(grep("love", twitter_data))/length(grep("hate", twitter_data))
sprintf("We get around: %s", love_hate)
```

The one tweet in the `en_US twitter data set` that matches the word `"biostats"` says what?
```{r}
tweet<-grep("biostats", twitter_data, value=TRUE)
tweet
```


How many tweets have the exact characters `"A computer once beat me at chess, but it was no match for me at kickboxing"`. (I.e. the line matches those characters exactly.)
```{r}
tweet_match<-grep("A computer once beat me at chess, but it was no match for me at kickboxing", twitter_data)
tweet_match
```

# *Task 2 - Exploratory Data Analysis* & **Task 3 - Modeling**
The first step in building a predictive model for text is understanding the distribution and relationship between the words, tokens, and phrases in the text. The goal of this task is to understand the basic relationships you observe in the data and prepare to build your first linguistic models.

_Tasks to accomplish_ 

1. **Exploratory analysis** perform a thorough exploratory analysis of the data, understanding the distribution of words and relationship between the words in the corpora. 
2. **Understand frequencies of words and word pairs** build figures and tables to understand variation in the frequencies of words and word pairs in the data.

_Questions to consider_

1. Some words are more frequent than others - what are the distributions of word frequencies? 
2. What are the frequencies of 2-grams and 3-grams in the dataset? 
3. How many unique words do you need in a frequency sorted dictionary to cover 50% of all word instances in the language? 90%? 
4. How do you evaluate how many of the words come from foreign languages? 
5. Can you think of a way to increase the coverage -- identifying words that may not be in the corpora or using a smaller number of words in the dictionary to cover the same number of phrases?

_Questions to consider_

1. How can you efficiently store an n-gram model (think Markov Chains)?
2. How can you use the knowledge about word frequencies to make your model smaller and more efficient?
3. How many parameters do you need (i.e. how big is n in your n-gram model)?
4. Can you think of simple ways to "smooth" the probabilities (think about giving all n-grams a non-zero probability even if they aren't observed in the data) ?
5. How do you evaluate whether your model is any good?
6. How can you use backoff models to estimate the probability of unobserved n-grams?



Using _Tidy Data_ methodology we can perform some EDA. We can use a simple Word Cloud to visualize the most frequent words.

```{r}
blogsmp <- sample(blog_data,1000)
newsmp <- sample(news_data,1000)
twittersmp <- sample(twitter_data,1000)
sample <- c(blogsmp,newsmp,twittersmp)
txt <- sent_detect(sample)
remove(blogsmp,newsmp,twittersmp,blog_data,news_data,twitter_data,sample)
```


Now we remove what we won't be needing
```{r}
txt <- removeNumbers(txt)
txt <- removePunctuation(txt)
txt <- stripWhitespace(txt)
txt <- tolower(txt)
txt <- txt[which(txt!="")]
txt <- data.frame(txt,stringsAsFactors = FALSE)
```

We now create some word clouds to visualize the most frequent words and order data frames of 1-grams, 2-grams, 3-grams

```{r}
words<-WordTokenizer(txt) 
grams<-NGramTokenizer(txt)
quadgram <- NGramTokenizer(corpus, Weka_control(min = 4, max = 4,delimiters = " \\r\\n\\t.,;:\"()?!"))

for(i in 1:length(grams)) 
{if(length(WordTokenizer(grams[i]))==2) break}
for(j in 1:length(grams)) 
{if(length(WordTokenizer(grams[j]))==1) break}

onegrams <- data.frame(table(words))
onegrams <- onegrams[order(onegrams$Freq, decreasing = TRUE),]
bigrams <- data.frame(table(grams[i:(j-1)]))
bigrams <- bigrams[order(bigrams$Freq, decreasing = TRUE),]
trigrams <- data.frame(table(grams[1:(i-1)]))
trigrams <- trigrams[order(trigrams$Freq, decreasing = TRUE),]
quadgram <- data.frame(table(quadgram))
quadgram <- quadgram[order(quadgram$Freq,decreasing = TRUE),]
remove(i,j,grams,quadgram)
```

### _Word cloud_
```{r}
library(wordcloud)
wordcloud(words, scale=c(5,0.1), max.words=100, random.order=FALSE, 
          rot.per=0.5, use.r.layout=FALSE, colors=brewer.pal(8,"Accent"))
```

```{r}
wordcloud(onegrams$words, onegrams$Freq, scale=c(5,0.5), max.words=300, random.order=FALSE, 
          rot.per=0.5, use.r.layout=FALSE, colors=brewer.pal(8,"Accent"))
```
The first graph shows the distribution of words in the corpora except such words, as "the", "a", "of", "to", etc. The second graph - the distribution of all single wors. The frequences lay between 3796 to 1.

What are the frequencies of 2-grams and 3-grams in the dataset?

## _Some Barplot_
```{r}
barplot(bigrams[1:20,2],col="plum3",
        names.arg = bigrams$Var1[1:20],srt = 45,
        space=0.1, xlim=c(0,20),las=2)
```

```{r}
barplot(trigrams[1:20,2],col="plum3",
        names.arg = trigrams$Var1[1:20],srt = 45,
        space=0.1, xlim=c(0,20),las=2)
```



### *Future work*

The goal is to create a predictive model which predicts the most probable words to follow an input from the user. This model will be evaluated and deployed as a shiny application.
